---
title: "analysis"
author: "Blublu"
date: "2025-04-03"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(effsize)
library(lmtest)
library(sandwich)
library(performance)
```


Load data
```{r}
df <- read.csv("../data/processed/data.csv", header = TRUE, row.names = 1)

df <- tibble(df)
df$cond <- as.factor(df$cond)
df_parent <- df %>% filter(gen != 0)
```


# Writing

## Task Performance

overall performance
```{r}
t.test(
  df %>% filter(cond == "ADV") %>% pull(score_corrected),
  df %>% filter(cond == "OBS") %>% pull(score_corrected),
  method = "s"
)

sd(df %>% filter(cond == "OBS") %>% pull(score_corrected))
sd(df %>% filter(cond == "ADV") %>% pull(score_corrected))

cohen.d(
  df %>% filter(cond == "ADV") %>% pull(score_corrected),
  df %>% filter(cond == "OBS") %>% pull(score_corrected)
)
```

first 3 vs last 3 performance
```{r}
first_gen <- "First 3 gen"
last_gen <- "Last 3 gen"

# Define groups
df_stat <- df %>%
  mutate(group = case_when(
    gen <= 2 ~ first_gen,
    gen >= 7 ~ last_gen,
  ))

# Function to perform t-test and compute standard deviation
compute_stats <- function(df, condition, group_var) {
  subset <- df %>% filter(cond == condition, group == group_var) %>% pull(score_corrected)
  list(
    t_test = subset,
    sd = sd(subset)
  )
}

# Perform t-tests and compute standard deviations
adv_first_gen_stats <- compute_stats(df_stat, "ADV", first_gen)
adv_last_gen_stats <- compute_stats(df_stat, "ADV", last_gen)

obs_first_gen_stats <- compute_stats(df_stat, "OBS", first_gen)
obs_last_gen_stats <- compute_stats(df_stat, "OBS", last_gen)

# Run t-tests
t.test(adv_first_gen_stats$t_test, adv_last_gen_stats$t_test, method = "s")
adv_first_gen_stats$sd
adv_last_gen_stats$sd
cohen.d(adv_first_gen_stats$t_test, adv_last_gen_stats$t_test)

t.test(obs_first_gen_stats$t_test, obs_last_gen_stats$t_test, method = "s", var.equal = TRUE)
obs_first_gen_stats$sd
obs_last_gen_stats$sd
cohen.d(obs_first_gen_stats$t_test, obs_last_gen_stats$t_test)

t.test(adv_first_gen_stats$t_test, obs_first_gen_stats$t_test, method = "s")
cohen.d(adv_first_gen_stats$t_test, obs_first_gen_stats$t_test)
t.test(adv_last_gen_stats$t_test, obs_last_gen_stats$t_test, method = "s")
cohen.d(adv_last_gen_stats$t_test, obs_last_gen_stats$t_test)
```


lm
```{r}
res <- lm(score_corrected ~ gen * cond, data = df)
summary(res)
```





## Model Knowledge

overall model knowledge
```{r}
t.test(
  df %>% filter(cond == "ADV") %>% pull(pl2_understanding),
  df %>% filter(cond == "OBS") %>% pull(pl2_understanding),
  method = "s"
)

sd(df %>% filter(cond == "OBS") %>% pull(pl2_understanding))
sd(df %>% filter(cond == "ADV") %>% pull(pl2_understanding))

cohen.d( 
  df %>% filter(cond == "ADV") %>% pull(pl2_understanding),
  df %>% filter(cond == "OBS") %>% pull(pl2_understanding)
)
```

first 3 vs last 3 model knowledge
```{r}
first_gen <- "First 3 gen"
last_gen <- "Last 3 gen"

# Define groups
df_stat <- df %>%
  mutate(group = case_when(
    gen <= 2 ~ first_gen,
    gen >= 7 ~ last_gen,
  ))

# Function to perform t-test and compute standard deviation
compute_stats <- function(df, condition, group_var) {
  subset <- df %>% filter(cond == condition, group == group_var) %>% pull(pl2_understanding)
  list(
    t_test = subset,
    sd = sd(subset)
  )
}

# Perform t-tests and compute standard deviations
adv_first_gen_stats <- compute_stats(df_stat, "ADV", first_gen)
adv_last_gen_stats <- compute_stats(df_stat, "ADV", last_gen)

obs_first_gen_stats <- compute_stats(df_stat, "OBS", first_gen)
obs_last_gen_stats <- compute_stats(df_stat, "OBS", last_gen)

# Run t-tests
t.test(adv_first_gen_stats$t_test, adv_last_gen_stats$t_test, method = "s")
adv_first_gen_stats$sd
adv_last_gen_stats$sd
cohen.d(adv_first_gen_stats$t_test, adv_last_gen_stats$t_test)

t.test(obs_first_gen_stats$t_test, obs_last_gen_stats$t_test, method = "s", var.equal = TRUE)
obs_first_gen_stats$sd
obs_last_gen_stats$sd
cohen.d(obs_first_gen_stats$t_test, obs_last_gen_stats$t_test)

t.test(adv_first_gen_stats$t_test, obs_first_gen_stats$t_test, method = "s")
cohen.d(adv_first_gen_stats$t_test, obs_first_gen_stats$t_test)
t.test(adv_last_gen_stats$t_test, obs_last_gen_stats$t_test, method = "s")
cohen.d(adv_last_gen_stats$t_test, obs_last_gen_stats$t_test)
```


lm
```{r}
res <- lm(pl2_understanding ~ gen * cond, data = df)
summary(res)
```






## Correlation Task Performance Model Knowledge


PARENT MODEL KNOWLEDGE
```{r}
res <- lm(score_corrected ~ parent_pl2_understanding * cond, data = df_parent)
summary(res)
confint(res)
```


MODEL KNOWLEDGE
```{r}
res <- lm(score_corrected ~ pl2_understanding * cond, data = df)
summary(res)
confint(res)
```


MODEL DIAGNOSTIC PARENT
```{r}
model <- lm(score_corrected ~ parent_pl2_understanding * cond, data = df_parent)

check_model(model)
```


PARENT TASK PERFORMANCE
```{r}
res <- lm(score_corrected ~ parent_score_corrected * cond, data = df_parent)
summary(res)
confint(res)
```



MODEL DIAGNOSTIC PARENT CHILDREN
```{r}
model <- lm(score_corrected ~ pl2_understanding * cond, data = df)

check_model(model)
```



## NLP

```{r}
df_NLP = read.csv("../old/NLP/results/docs_topic_weights.csv")
df_NLP <- df_NLP %>%
  mutate(gen = gen -1)
```


```{r}
data <- df_NLP %>%
  filter(gen != 0) %>%
  select(
    ID,
    pl2_understanding,
    gen,
    parent_topic_0,
    parent_topic_1,
    parent_topic_2,
    parent_topic_3,
    parent_topic_4,
    parent_topic_5,
    parent_topic_6,
    parent_topic_7,
    parent_topic_8
  ) %>%
  gather(key = "topic_n", value = "prop", -c(ID, pl2_understanding, gen)) %>%
  mutate(topic_n = factor(topic_n))

model <- lmer(pl2_understanding ~ prop:topic_n + (1 | gen), data = data)
summary(model)
```


perf
```{r}
data <- df_NLP %>%
  filter(gen != 0) %>%
  select(
    ID, score_corrected, gen,
    parent_topic_0,
    parent_topic_1,
    parent_topic_2,
    parent_topic_3,
    parent_topic_4,
    parent_topic_5,
    parent_topic_6,
    parent_topic_7,
    parent_topic_8
  ) %>%
  gather(key = "topic_n", value = "prop", -c(ID, score_corrected, gen)) %>%
  mutate(topic_n = factor(topic_n))

model <- lmer(score_corrected ~ prop:topic_n + (1 | gen), data = data)
summary(model)
```


topic 2 growth lm
```{r}
res <- lm(topic_2 ~ gen, data = df_NLP)
summary(res)
```


topic 2 growth glm + betareg (ChatGPT)
```{r}
library(betareg)

df_betareg <- df_NLP %>%
  mutate(assigned_topic = factor(assigned_topic))

model_topic2 <- betareg(topic_2 ~ gen, data = df_betareg)

summary(model_topic2)


df_betareg$assigned_topic_2 <- as.integer(df_betareg$assigned_topic == 2)

model_assigned_topic2 <- glm(assigned_topic_2 ~ gen, family = binomial, data = df_betareg)

summary(model_assigned_topic2)
```


```{r}
df_topic2 <- df_NLP %>%
  filter(assigned_topic == 2)
df_not_topic2 <- df_NLP %>%
  filter(assigned_topic != 2)

t.test(df_topic2$score_corrected, df_not_topic2$score_corrected)
```


get parent_assigned_topic
```{r}
parent_assigned_topic <- rep(NA, nrow(df_NLP)) # Pre-fill with NAs

for (i in 1:nrow(df_NLP)){
  parent_ID <- df_NLP[i,]$parent_ID
  parent <- df_NLP %>% filter(ID == !!parent_ID) 
  parent_assigned_topic[i] <- ifelse(nrow(parent) > 0, parent$assigned_topic, NA)
}

df_NLP <- df_NLP %>%
  mutate(parent_assigned_topic = parent_assigned_topic)
```


```{r}
df_NLP %>%
  filter(gen != 0) %>%
  filter(assigned_topic == 2) %>%
  ggplot( aes(x = factor(parent_assigned_topic))) +
  geom_bar(fill = "steelblue", color = "black") +
  facet_wrap(~ gen) +
  labs(title = "Distribution of parent_assigned_topic in participant with topic 2",
       x = "Parent Assigned Topic",
       y = "Count") +
  theme_minimal()
```


```{r}
df_NLP %>%
  filter(assigned_topic == 2) %>%
  count(parent_assigned_topic) %>%
  arrange(desc(n))

100 / nrow(df_NLP %>% filter(assigned_topic == 2))
```


search if participant had children (is_parent)
```{r}
is_parent <- rep(NA, nrow(df_NLP)) # Pre-fill with NAs
n_children <- rep(NA, nrow(df_NLP))

for (i in 1:nrow(df_NLP)){
  ID <- df_NLP[i, ]$ID
  children <- df_NLP %>% filter(parent_ID == !!ID)
  is_parent[i] <- ifelse(nrow(children) > 0, TRUE, FALSE)
  n_children[i] <- nrow(children)
}

df_NLP <- df_NLP %>%
  mutate(is_parent = is_parent) %>%
  mutate(n_children = n_children)
``` 


```{r}
model <- glm(is_parent ~ factor(assigned_topic) * factor(gen), data = df_NLP, family = binomial)
summary(model)
```





