---
title: "resulting_advice"
author: "Blublu"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(reshape2)
library(colorspace)
```


```{r}
df <- read.csv("../output/processed/data_stake.csv") %>% select(-X)
df %>% filter(gen == 10) %>% filter(rank <= 10) %>% pull(ADVICE)
```


Using the above advice, I prompted:
" This is a database of written advice (10):

Using only the information in these advice, write yourself what you would consider to be the best advice possible. Your advice should follow the same structure (multiple sentences, no bullet point) than the others. "

And got this answer:

There are two types of baskets: RED and YELLOW. RED baskets are associated with gnomes of the following colours: RED, PINK, PURPLE, and BROWN, while YELLOW baskets are linked to YELLOW, ORANGE, BLUE, and GREEN gnomes. To maximize your rewards, choose gnomes leading to one basket type until the number of mushrooms drops to 0, 1, or 2, then switch to the other basket. Stick with the new basket until its rewards also decline, then switch back. At times, both baskets may produce low results; in these cases, stay with one basket for a few rounds before switching again. This back-and-forth approach ensures you stay aligned with changes in basket values. Keep an eye on mushroom patterns, as they may shift randomly, requiring you to adapt your strategy throughout the game. Good luck!


```{r}
gpt_advice <- "There are two types of baskets: RED and YELLOW. RED baskets are associated with gnomes of the following colours: RED, PINK, PURPLE, and BROWN, while YELLOW baskets are linked to YELLOW, ORANGE, BLUE, and GREEN gnomes. To maximize your rewards, choose gnomes leading to one basket type until the number of mushrooms drops to 0, 1, or 2, then switch to the other basket. Stick with the new basket until its rewards also decline, then switch back. At times, both baskets may produce low results; in these cases, stay with one basket for a few rounds before switching again. This back-and-forth approach ensures you stay aligned with changes in basket values. Keep an eye on mushroom patterns, as they may shift randomly, requiring you to adapt your strategy throughout the game. Good luck!"
```


AFTER PYTHON SIMILARITY
```{r}
df <- read.csv("../output/processed/similarity_matrix_perfect_advice.csv", header = TRUE)

df_similarity_gen <- df %>%
  group_by(gen) %>%
  summarise(
    similarity = mean(similarity)
  )

color_palette <- c("#FFEDA0", "#FED976", "#FECC5A", "#FEB24C", "#FD8D3C", "#FC4E2A", "#E31A1C", "#BD0026", "#800026", "#4C001A")

plot <- df_similarity_gen %>%
  mutate(gen = as.factor(gen)) %>%
  ggplot(aes(x = gen, y = similarity)) +
  
  geom_col(
    aes(color = gen, fill = after_scale(lighten(color, 0.5))),
    linewidth = 1.2
  ) +

  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette, guide = "none") +
  scale_y_continuous(expand = c(.001, .001)) +
  
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line = element_line(colour = "black", size = .8),
    axis.ticks = element_line(colour = "black", size = .8),
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 8), size = 12),
    axis.title.y = element_text(margin = margin(r = 8), size = 12)
  ) +
  
  labs(
    title = "Similarity within generation to 'perfect' advice",
    x = "Generation",
    y = "Similarity"
  )

plot

ggsave("../figures/test/similarity_with_perfect_advice.tiff", bg = "white")
# ggsave("../figures/test/fig6_similarity_with_perfect_advice.svg", bg = "white", device = "svg")
```

violin plot similarity resulting advice
```{r}
df <- read.csv("../output/processed/similarity_matrix_perfect_advice.csv", header = TRUE)

median_values <- df %>%
  group_by(gen) %>%
  summarize(median_score = median(similarity, na.rm = TRUE)) %>%
  mutate(gen = as.numeric(as.character(gen))) # Convert gen to numeric if necessary

max_value <- max(df$similarity, na.rm = TRUE)
min_value <- min(df$similarity, na.rm = TRUE)
min_value <- floor(min_value * 10) / 10

plot <- df %>%
  mutate(gen = as.factor(gen)) %>%
  ggplot(aes(x = gen, y = similarity)) +
  ggdist::stat_halfeye(
    fill = "#870052",
    adjust = .5,
    width = .4,
    .width = 0,
    justification = -.3,
    point_colour = NA
  ) +
  geom_boxplot(
    color = "#870052",
    fill = "#E5A6C9",
    width = 0.15,
    outlier.shape = NA
  ) +
  geom_point(
    color = "#870052",
    fill = "#E5A6C9",
    size = 1.5,
    alpha = .2,
    position = position_jitter(
      seed = 1, width = .02
    )
  ) +
  geom_line(
    data = median_values,
    aes(x = gen, y = median_score),
    group = 1,
    color = "black",
    linewidth = 1.5
  ) +
  geom_point(
    data = median_values,
    aes(x = gen, y = median_score),
    color = "black",
    size = 3
  ) +
  labs(
    x = "Generation",
    y = "Similarity"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line = element_line(colour = "black", size = .8),
    axis.ticks = element_line(colour = "black", size = .8),
    axis.ticks.length.x = unit(5, "pt"),
    axis.ticks.length.y = unit(7, "pt"),
    axis.text = element_text(family = "Roboto Mono", size = 11, colour = "black"),
    axis.title.x = element_text(
      margin = margin(t = 10),
      size = 16
    ),
    axis.title.y = element_text(
      margin = margin(r = 10),
      size = 16
    ),
    plot.margin = margin(15, 15, 10, 15),
    legend.position = "none"
  )

ggsave("../figures/test/similarity_with_perfect_advice_violin.tiff", bg = "white")
```

# HEATMAP TEST
```{r}
df <- read.csv("../output/processed/data_stake.csv") %>% 
  select(-X)
df_similarity_matrix <- read.csv("../output/processed/similarity_matrix_perfect_advice.csv", header = TRUE) %>%
  select(-gen) %>%
  filter(model == "distiluse-base-multilingual-cased-v1")
```


```{r}
df_heatmap <- left_join(df, df_similarity_matrix, by = "ID")
```


```{r}
library(viridis)

df_heatmap <- left_join(df, df_similarity_matrix, by = "ID")

df_heatmap <- df_heatmap %>%
  select(
    ID,
    gen,
    score_corrected,
    similarity,
    ADVICE
  ) %>%
  group_by(gen) %>%
  mutate(rank = rank(score_corrected, ties.method = "first"))


df_heatmap %>%
  ggplot( aes(x = as.factor(rank), y = as.factor(gen), fill = similarity)) +
  geom_tile() +
  scale_fill_viridis(discrete = FALSE) +
  labs(
    x = "Performance",
    y = "Generation"
  ) +
  scale_x_discrete(
    breaks = c("1", "100"),
    labels = c("Worst", "Best")
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.text = element_text(family = "Roboto Mono", size = 11, colour = "black"),
    axis.title.x = element_text(
      margin = margin(t = 10),
      size = 16
    ),
    axis.title.y = element_text(
      margin = margin(r = 10),
      size = 16
    ),
    plot.margin = margin(15, 15, 10, 15)
  )

df_heatmap %>%
  group_by(gen) %>%
  mutate(similarity_rank = rank(similarity, ties.method = "first")) %>%
  ggplot( aes(x = as.factor(similarity_rank), y = as.factor(gen), fill = similarity)) +
  geom_tile() +
  scale_fill_viridis(discrete = FALSE) +
  labs(
    x = "Similarity",
    y = "Generation"
  ) +
  scale_x_discrete(
    breaks = c("1", "100"),
    labels = c("Least similar", "Most similar")
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.text = element_text(family = "Roboto Mono", size = 11, colour = "black"),
    axis.title.x = element_text(
      margin = margin(t = 10),
      size = 16
    ),
    axis.title.y = element_text(
      margin = margin(r = 10),
      size = 16
    ),
    plot.margin = margin(15, 15, 10, 15)
  )
```
```{r}
df_fitness <- df %>%
  select(
    ID,
    gen,
    score_corrected,
    pl1_understanding,
    pl2_understanding,
    rank
  )

i <- 1

for (id in df$ID) {
  df_children <- df %>% filter(parent_ID == id)
  n_C <- nrow(df_children)
  n_GC <- 0
  for (child_id in df_children$ID) {
    df_grandchildren <- df %>% filter(parent_ID == child_id)
    n <- nrow(df_grandchildren)
    n_GC <- n_GC + n
  }
  df_fitness[i, "n_C"] <- n_C
  df_fitness[i, "n_GC"] <- n_GC
  i <- i + 1
}
```



```{r}
library(viridis)
library(scales)

df_heatmap <- left_join(df, df_similarity_matrix, by = "ID")
df_heatmap <- left_join(df_heatmap, df_fitness %>% select(ID, n_GC, n_C), by = "ID")

df_heatmap <- df_heatmap %>%
  select(
    ID,
    gen,
    n_GC,
    n_C,
    score_corrected,
    similarity,
    ADVICE
  ) %>%
  group_by(gen) %>%
  mutate(rank = rank(score_corrected, ties.method = "first")) %>%
  mutate(fitness = ifelse(gen < 9, n_GC, 0)) %>%
  ungroup() %>%
  mutate(fitness = rescale(fitness)) %>%
  mutate(fitness = ifelse(gen >= 9, .5, fitness))


ggplot(df_heatmap, aes(x = as.factor(rank), y = as.factor(gen), col = similarity)) +
  geom_point(aes(size = fitness), shape = 15)




df_plot <- tibble(
  x = rep(1:100, 10),
  y = rep(1:10, each = 100),
  widths = rep(10, 1000)
)

ggplot(df_plot, aes(x = as.factor(x), y = as.factor(y))) +
  geom_tile(
    aes(width = widths),
    fill = "white",
    color = "black"
  ) + 
  # scale_x_continuous(expand = c(0, 0)) +
  # scale_y_continuous(expand = c(0, 0)) +
  # coord_fixed(ratio = 1) + # Fix the aspect ratio to make tiles square
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove grid lines
    axis.title = element_blank()  # Remove axis titles
  )

```





```{r}
library(viridis)
df_heatmap <- df_heatmap %>%
  select(
    ID,
    gen,
    pl2_understanding,
    similarity,
    ADVICE
  ) %>%
  group_by(gen) %>%
  mutate(rank = rank(pl2_understanding, ties.method = "first"))


df_heatmap %>%
  ggplot( aes(x = as.factor(rank), y = as.factor(gen), fill = similarity)) +
  geom_tile() +
  scale_fill_viridis(discrete = FALSE) +
  labs(
    x = "Transition knowledge",
    y = "Generation"
  ) +
  scale_x_discrete(
    breaks = c("1", "100"),
    labels = c("Worst", "Best")
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.text = element_text(family = "Roboto Mono", size = 11, colour = "black"),
    axis.title.x = element_text(
      margin = margin(t = 10),
      size = 16
    ),
    axis.title.y = element_text(
      margin = margin(r = 10),
      size = 16
    ),
    plot.margin = margin(15, 15, 10, 15)
  )
```


```{r}
library(viridis)

df_heatmap <- left_join(df, df_similarity_matrix, by = "ID")
df_heatmap <- df_heatmap %>%
  select(
    ID,
    gen,
    pl1_understanding,
    similarity,
    ADVICE
  ) %>%
  group_by(gen) %>%
  mutate(rank = rank(pl1_understanding, ties.method = "first"))


df_heatmap %>%
  ggplot( aes(x = as.factor(rank), y = as.factor(gen), fill = similarity)) +
  geom_tile() +
  scale_fill_viridis(discrete = FALSE) +
  labs(
    x = "Pair knowledge",
    y = "Generation"
  ) +
  scale_x_discrete(
    breaks = c("1", "100"),
    labels = c("Worst", "Best")
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.text = element_text(family = "Roboto Mono", size = 11, colour = "black"),
    axis.title.x = element_text(
      margin = margin(t = 10),
      size = 16
    ),
    axis.title.y = element_text(
      margin = margin(r = 10),
      size = 16
    ),
    plot.margin = margin(15, 15, 10, 15)
  )
```





